---
title: 'GY401W: Invertebrate paleontology'
date: "Spring 2024"
output:
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    latex_engine: pdflatex
    template: Syllabus_template.tex
  html_document:
    df_print: paged
geometry: margin=1in

#Fill this in with any number of instructors. It uses a two column layout with a column break after each instructor.

instructor:
- name: Dr. Benjamin J. Linzmeier
  email: blinzmeier@southalabama.edu
  prefered_contact: Canvas LMS messages
  officehours: 10:00 am - 12:00 pm MW \newline 1:00 - 3:00 pm T
  office: Life Science Building room 344
  correspondence_policy: Within 24 hours on weekdays; 48 hours on weekends

# You can include multiple entries for a teaching assistant by filling in all these options below. Currently this is set to populate two columns on the syllabus but there is not an even column break inserted.

# teachingassistant:
#  - name: Filler
#    email: filler@fill.com
#    officehours: W T
#    office: OfficeNumber
#    sections: '301'
   
classroom: 'ELCSB 0119	'
classhours: 9:30 - 10:45 am	TTh
#labroom: 'LCSB 337'
#labhours: 1:25 pm - 3:55 pm	W
fontfamily: mathpazo
fontsize: 11pt
header-includes: \linespread{1.05}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(timeDate)
library(tibble)
library(dplyr)
library(knitr)

#This code chunk sets up a table of meeting days for the course.
#You can set days of the week for course meetings (e.g. Monday and Wednesday).
#Semester start and end dates are selected and then university holidays like spring break are specified.

#Days of the week that the course meets. Must be full day name in quotes.
Days <- c("Tuesday", "Thursday")

#Semester start date for classes as YYYY/M/D.
SemesterStart <- as.Date("2023/8/22")

#Semester end date for classes as YYYY/M/D.
SemesterEnd <- as.Date("2023/12/8")

#Fall break start and stop date.
FallBreak <-
  seq.Date(
    from = as.Date("2023/10/12"),
    to = as.Date("2023/10/15"),
    by = "day"
  )

#Spring break start and stop date.
SpringBreak <-
  seq.Date(
    from = as.Date("2024/03/04"),
    to = as.Date("2024/03/10"),
    by = "day"
  )

#Creates a list of days within the semester.
SemesterDays <-
  seq.Date(from = SemesterStart, to = SemesterEnd, by = "day")

#Add day before Thanksgiving.
DayBeforeThnks<-as.Date(holiday(
    year = getRmetricsOptions("currentYear"),
    Holiday = c("USThanksgivingDay")))-1

#Full list of holidays that there are no classes including breaks.
Holidays <-
  c(as.Date(holiday(
    year = getRmetricsOptions("currentYear"),
    Holiday = c("USLaborDay", "USThanksgivingDay", "USMLKingsBirthday")
  )), SpringBreak, FallBreak, DayBeforeThnks)

#Create a list of semester days excluding holidays.
SemesterDays <- SemesterDays[!SemesterDays %in% Holidays]

#Make dataframe with all semester days.
ClassDays <- data.frame(index = 1:length(SemesterDays))

#Insert dates.
ClassDays$Date <- SemesterDays

#Insert weekday designation.
ClassDays$Weekday <- weekdays(ClassDays$Date)

#Limit to the class meeting days.
MeetingDays <- ClassDays[ClassDays$Weekday %in% Days, ]

#Create numeric list of weeks.
Week <- cut.Date(MeetingDays$Date, breaks = "week")
levels(Week) <- 1:length(levels(Week))

MeetingDays$Week <- Week
MeetingDays$Class <- 1:nrow(MeetingDays)

#Replace this with c("Lecture topic 1",...,"Lecture topic N") to fill in table.

Lecture_Topic <- c("Introduction",
                   "Earth History as Science",
                   "The Universe",
                   "The Earth, broadly",
                   "Hadean and Archean",
                   "Archean Eon",
                   "Evolution 1",
                   "Evolution 2",
                   "Proterozoic Geology 1",
                   "Proterozoic Geology 2",
                   "Proterozoic Life 1",
                   "Proterozoic Life 2",
                   "Common Cause",
                   "Paleozoic Geology 1",
                   "Midterm Review",
                   "Midterm",
                   "Paleozoic Geology 2",
                   "Paleozoic Life 1",
                   "Paleozoic Life 2",
                   "Mesozoic Geology 1",
                   "Mesozoic Geology 2",
                   "Mesozoic Life 1",
                   "Mesozoic Life 2",
                   "Cenozoic Geology 1",
                   "Cenozoic Geology 2",
                   "Cenozoic Life 1",
                   "Cenozoic Life 2",
                   "Models in Earth Science",
                   "Climate and the future 1",
                   "Climate and the future 2")
  #1:nrow(MeetingDays)

Reading_Assignment <- c("Introduction",
                        "",
                        "Birth",
                        "",
                        "Blue Earth",
                        "",
                        "Living Earth",
                        "",
                        "Red Earth",
                        "",
                        "The 'Boring' Billion",
                        "",
                        "White Earth",
                        "",
                        "",
                        "TBD",
                        "",
                        "Green Earth",
                        "",
                        "TBD",
                        "",
                        "TBD",
                        "",
                        "TBD",
                        "",
                        "TBD",
                        "",
                        "TBD",
                        "The Future",
                        "")

MeetingDays <- bind_cols(MeetingDays, Lecture = Lecture_Topic, Reading = Reading_Assignment)

#### Grade table ####
Grades <- tribble( ~ Grade,
                   ~ Range,
                   "A",
                   "90-100 %",
                   "B",
                   "80-89 %",
                   "C",
                   "70-79 %",
                   "D",
                   "60-69 %",
                   "F",
                   "< 60 %")

####Assignment weight table####
Weights <- tribble(
                    ~ Item,
                    ~ Weight,
                    ~ Graded,
                    "Discussion posts",
                    "25 %",
                    "Effort",
                    "Gots and nots",
                    "10 %",
                    "Completion",
                    "Tophat questions",
                    "15 %",
                    "Completion (80%)/Correct (20%)",
                    "Weekly Quizzes",
                    "25 %",
                    "Correct",
                    "Midterm Exam",
                    "10 %",
                    "Correct",
                    "Final Exam",
                    "15 %",
                    "Correct"
                  )


```

# Course description

What can studying the history of Earth tell us about the future and the present? In this course we will walk though major events in the history of our planet and dive into the processes and peculiarities that caused them. We will connect these events to places that matter to us and learn about how their occurrence shapes life today. By the end of the semester you will have a broad understanding of the science of Geology and its application to studying how the Earth works.

# Important Dates
August 29th - Last day to drop without WD\
October 10th - Midterm\
November 3rd - Last day to drop or withdraw.
See the [Academic Calendar](https://www.southalabama.edu/academiccalendar/) for more information.\
December 14th - Final 10:30 am - 12:30 pm\

# Course goals

1.    Explain and apply the scientific method as it is practiced in the Geosciences.
2.    Link observations of the Earth on multiple scales to the development of theories.
3.    Identify common invertebrate fossils.
4.    Describe a broad history of the Earth.
5.    Link Geological History to current landscapes, biology, and climate.


# Class textbook

We will be reading a book written for a general audience this semester. It has fewer figures than a textbook, but a better narrative. It is also less expensive.

Hazen, Robert M. __The Story of Earth: The First 4.5 Billion Years, from Stardust to Living Planet.__ Reprint edition. Penguin Books, 2013.

### Lectures and papers
Both lecture slides and supplemental readings will be available via Canvas. Slides will generally be posted shortly before lecture.

# Tophat

We will be using Tophat to facilitate active learning during class. It is relatively inexpensive and can be used in other classes as well. If you encounter a hardship in activating it, please let me know as soon as possible.

# Canvas

We will be using Canvas for quizzes, discussions, and your gots-and-nots. These will be organized into weekly modules. New discussion posts will be available on Saturday and quizzes for the previous week will be due on Saturday. Quizzes will be accessible Thursday afternoon.

\newpage

```{r Schedule, echo=FALSE}
kable(MeetingDays[,c("Date",
                     "Class",
                     "Lecture",
                     "Reading")
                     ],
  caption = "Course Schedule.",
  align = c("l",
            "c",
            "c",
            "r")
)
```
\newpage

# Student conduct

## Attendance policy

Students who complete less than 50% of assignments/exams will be assigned a failing (F) grade. Review the 'Attendance and Absences Policy', in the Undergraduate and Graduate Bulletin for attendance and absences policy.

Lectures can be attened in person or via Zoom. Recordings will be made available by request.

## Grading

### Late policy
I will be grading assignments in batches to provide fair assessment to everyone. This means late work can be disruptive. The penalty for late work will be 20% a day after the first 24 hours late. If you have a reason (illness, etc.) the late policy will be relaxed.

### Dropped grades
Two weeks worth of your lowest scores will be dropped at the end of the semester. This means that the lowest two quiz scores, discussion posts, and Tophat questions will be removed when calculating your final grade.

### Extra credit
Extra credit may be made available throughout the semester at my discretion.

Assignments will be returned within a week of submission.

```{r Grades_Table, echo=FALSE}
kable(Grades,
      align = c("l", "r"))
```

Assignment weighting will follow:
```{r Weights_Table, echo=FALSE}
kable(Weights,
      align = c("l", "r"))
```

<!-- The sections of code below that include "child=..." are drawing from seperate .md documents. These allow for updating text in a single spot for use in multiple syllabus documents. -->
```{r Academic disruption, child='general_components/academic_disruption.md'}

```

```{r Academic honesty, child='general_components/academic_honesty.md'}

```

## Course and Teacher evaluation
Student input for the purpose of course improvement is taken very seriously and will potentially be done periodically. Please take the time to evaluate this course and the instructor, especially at the end of the semester. Evaluations will in no way affect your grade.

# Safety

## Building Safety

### Fire
Random fire safety drills may occur during the course of the term. You will be expected to evacuate the building and assemble at the designated location: outside by the northeast corner of the building (follow your instructor).

### Tornado
Move away from external windows to the hallway in the basement, first, or second floor. 

### Active Shooter Incident
"Get Out" of the building as long as they can do so safely. Once you have exited the building and are at a safe location, "Call Out" to law enforcement using 911, USA Police at 460-6312, or press the Emergency button on your LiveSafe Mobile Safety App. If you can't get out, "Shelter in Place" in a secure location that is away from windows and doors, preferably a location that can be locked and barricaded.


# Campus resources

```{r COVID policy, child='general_components/COVID_policy.md'}

```

```{r Writing Center, child='general_components/writing_center.md'}

```

```{r Disability Services, child='general_components/disability_services.md'}

```

```{r Food Insecurity, child='general_components/food_insecurity.md'}

```

```{r Mental Health, child='general_components/mental_health.md'}

```

```{r Disaster Plan, child='general_components/disaster_plan.md'}

```

# Final thoughts

This document is a roadmap for our semester. We learn about the Earth together and our individual experiences shape how we interpret and value data. Like all your classes, you will get out what you put into this course. Asking for help from one another and your instructors is important, don't be afraid to ask a question about something you don't know or if you want to check your knowledge about something you think you know.

__If this document is updated, a copy will be supplied to you via Canvas and changes will be announced in class.__