---
title: 'GY401W: Invertebrate paleontology'
date: "Spring 2024"
output:
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    latex_engine: pdflatex
    template: Syllabus_template.tex
  html_document:
    df_print: paged
geometry: margin=1in

#Fill this in with any number of instructors. It uses a two column layout with a column break after each instructor.

instructor:
- name: Dr. Benjamin J. Linzmeier
  email: blinzmeier@southalabama.edu
  prefered_contact: Canvas LMS messages
  officehours: 10:00 am - 12:00 pm MW \newline 1:00 - 3:00 pm T
  office: Life Science Building room 344
  correspondence_policy: Within 24 hours on weekdays; 48 hours on weekends

# You can include multiple entries for a teaching assistant by filling in all these options below. Currently this is set to populate two columns on the syllabus but there is not an even column break inserted.

# teachingassistant:
#  - name: Filler
#    email: filler@fill.com
#    officehours: W T
#    office: OfficeNumber
#    sections: '301'
   
classroom: 'ELCSB 337'
classhours: 12:20 - 1:10 pm	MWF
labroom: 'ELCSB 337'
labhours: 1:25 pm - 3:55 pm	W
fontfamily: mathpazo
fontsize: 11pt
header-includes: \linespread{1.05}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(timeDate)
library(tibble)
library(dplyr)
library(knitr)

#This code chunk sets up a table of meeting days for the course.
#You can set days of the week for course meetings (e.g. Monday and Wednesday).
#Semester start and end dates are selected and then university holidays like spring break are specified.

#Days of the week that the course meets. Must be full day name in quotes.
Days <- c("Monday", "Wednesday", "Friday")
#Semester start date for classes as YYYY/M/D.
SemesterStart <- as.Date("2024/1/9")

#Semester end date for classes as YYYY/M/D.
SemesterEnd <- as.Date("2024/4/28")

#Fall break start and stop date.
FallBreak <-
  seq.Date(
    from = as.Date("2023/10/12"),
    to = as.Date("2023/10/15"),
    by = "day"
  )

#Spring break start and stop date.
SpringBreak <-
  seq.Date(
    from = as.Date("2024/03/04"),
    to = as.Date("2024/03/10"),
    by = "day"
  )

#Creates a list of days within the semester.
SemesterDays <-
  seq.Date(from = SemesterStart, to = SemesterEnd, by = "day")

#Add day before Thanksgiving.
DayBeforeThnks<-as.Date(holiday(
    year = getRmetricsOptions("currentYear"),
    Holiday = c("USThanksgivingDay")))-1

#Full list of holidays that there are no classes including breaks.
Holidays <-
  c(as.Date(holiday(
    year = getRmetricsOptions("currentYear"),
    Holiday = c("USLaborDay", "USThanksgivingDay", "USMLKingsBirthday")
  )), SpringBreak, FallBreak, DayBeforeThnks)

#Create a list of semester days excluding holidays.
SemesterDays <- SemesterDays[!SemesterDays %in% Holidays]

#Make dataframe with all semester days.
ClassDays <- data.frame(index = 1:length(SemesterDays))

#Insert dates.
ClassDays$Date <- SemesterDays

#Insert weekday designation.
ClassDays$Weekday <- weekdays(ClassDays$Date)

#Limit to the class meeting days.
MeetingDays <- ClassDays[ClassDays$Weekday %in% Days, ]

#Create numeric list of weeks.
Week <- cut.Date(MeetingDays$Date, breaks = "week")
levels(Week) <- 1:length(levels(Week))

MeetingDays$Week <- Week
MeetingDays$Class <- 1:nrow(MeetingDays)

#Replace this with c("Lecture topic 1",...,"Lecture topic N") to fill in table.

Lecture_Topic <- c("Introduction and project outline",
                   "Writing Discussion",
                   "Fossilization",
                   "Sampling the record",
                   "Writing Discussion",
                   "Phanerozoic diversity",
                   "Phanerozoic diversity",
                   "Writing Discussion",
                   "Describing species",
                   "Phylogenetics",
                   "Writing Discussion",
                   "Species concepts",
                   "Describing variability",
                   "Writing Discussion",
                   "Growth and development",
                   "Allometric growth and ontogeny",
                   "Writing Discussion",
                   "Evolution and fossils",
                   "Mode and tempo of evolution",
                   "Writing Discussion",
                   "Midterm Review",
                   "Midterm",
                   "Writing Discussion",
                   "Functional morphology",
                   "Theoretical morphology",
                   "Writing Discussion",
                   "Global diversity",
                   "Sampling diversity",
                   "Writing Discussion",
                   "Mass extinctions",
                   "Recognizing mass extinctions",
                   "Writing Discussion",
                   "Cambrian explosion",
                   "Causes of the explosion",
                   "Writing Discussion",
                   "Late Permian Extinction",
                   "Determining causes of PT",
                   "Writing Discussion",
                   "End Cretaceous extinction",
                   "Causes of the K-Pg",
                   "Writing Discussion",
                   "Conservation paleobiology",
                   "Modern extinction?",
                   "Writing Discussion")
  #1:nrow(MeetingDays)

Reading_Assignment <- c("TBD")
# Syllabus and Project handout",
#                         "The Nature of the Fossil Record",
#                         "",
#                         "Alroy et al.,...?",
#                         "",
#                         "Systematics",
#                         "",
#                         "Populations and Species",
#                         "",
#                         "Growth and form",
#                         "",
#                         "Evolutionary rates and trends",
#                         "White Earth",
#                         "",
#                         "",
#                         "TBD",
#                         "",
#                         "Green Earth",
#                         "",
#                         "TBD",
#                         "",
#                         "TBD",
#                         "",
#                         "TBD",
#                         "",
#                         "TBD",
#                         "",
#                         "Putting the dead to work",
#                         "TBD")

MeetingDays <- bind_cols(MeetingDays, Lecture = Lecture_Topic, Reading = Reading_Assignment)

#### Grade table ####
Grades <- tribble( ~ Grade,
                   ~ Range,
                   "A",
                   "90-100 %",
                   "B",
                   "80-89 %",
                   "C",
                   "70-79 %",
                   "D",
                   "60-69 %",
                   "F",
                   "< 60 %")

####Assignment weight table####
Weights <- tribble(
                    ~ Item,
                    ~ Weight,
                    "Writing scaffolding",
                    "10 %",
                    "Semester Project",
                    "30 %",
                    "Labratory Assignments",
                    "30 %",
                    "Midterm Exam",
                    "10 %",
                    "Final Exam",
                    "20 %"
                  )


```

# Course description

How do fossils inform us about the history of life on Earth? In Paleontology (GY401W) we will work on linking change in common marine invertebrate fossils to change in environments and the mode and tempo of evolutionary change. Our focus will include a survey of modern paleontology research methods. Writing and conducting research will be an important component of the class.

# Writing requirements

For this class to fulfill the (W) requirements of USA, you will be doing a semester-long research project in an area related to paleontology. To build your skills in discipline-specific writing, we will be doing smaller writing assignments throughout the semester. The final projects will be peer-reviewed, revised, and then submitted for final grading and assessment.

# Important Dates
January 16th - Last day to drop without WD\
April 7th - Last day to drop or withdraw\
May ?? - Final exam at ???
See the [Academic Calendar](https://www.southalabama.edu/academiccalendar/) for more information.

\newpage

# Course goals

1.  Observe, describe, classify, and communicate characteristics of major fossil invertebrates.

2.  Develop hypotheses using fossil materials.

3.  Communicate paleontological research in writing to a variety of audiences.

4.  Read and summarize scientific papers focused on paleontology and related fields.

# Class textbooks

There is no required textbook for this spring, however I recommend getting these books from the library or used versions to compliment the lectures. Without a required textbook I will also be assigning technical papers to read throughout the semester. We will discuss these and you will build your technical reading skills along with your writing skills.

### For general Paleontology
Arnold I. Miller and Michael Foote, __Principles of Paleontology__, 5th edition or earlier edition, Pearson, ISBN: 9780321643186

### Field focused rock ID
Dorrik A.V. Stow, __Sedimentary Rocks in the Field: A Colour Guide__, Manson Publishing Ltd, ISBN 10: 1874545693

### Improve the mechanics of your writing
William Strunk JR. and E.B. White, __The Elements of Style__, ISBN: 1594200696

### Improve the process of writing
Stephen B. Heard, __The Scientists Guide to Writing__, ISBN: 1594200696

### Lectures and papers
Both lecture slides and papers will be available via Canvas. Slides will mostly be posted after lectures and assigned papers will be available in Canvas.


\newpage

```{r Schedule, echo=FALSE}
kable(MeetingDays[,c("Date",
                     "Class",
                     "Week",
                     "Lecture")
                     ],
  caption = "Course Schedule.",
  align = c("l",
            "c",
            "c",
            "l")
)
```

\newpage

# Student conduct

## Attendance policy

-   Students who complete less than 50% of assignments/exams will be assigned a failing (F) grade.

-   Review the 'Attendance and Absences Policy', in the Undergraduate and Graduate Bulletin for attendance and absences policy

## Grading

### Late policy
I will be grading assignments in batches to provide fair assessment to everyone. This means late work can be disruptive. The penalty for late work will be 20% a day after the first 24 hours late (0% possible after 5 days). If you have a reasonable excuse (illness, etc.) the late policy may be relaxed.

### Extra credit
Extra credit may be made available throughout the semester at my discretion.

Assignments will be returned within a week of submission.

```{r Grades_Table, echo=FALSE}
kable(Grades,
      align = c("l", "r"))
```

Assignment weighting will follow:
```{r Weights_Table, echo=FALSE}
kable(Weights,
      align = c("l", "r"))
```

\newpage
# Writing tasks
## Scaffolding tasks
Scaffolding writing tasks assigned to give you some practice and general feedback to improve your writing.

### Abstract 1

First attempt at writing an abstract using the Nature paragraph template.

### Figure Caption

Write the figure captions for several example figures using some literature examples. Learn to describe figures in text.

### Abstract 2

Write the abstract for a second paper using the feedback from the first abstract to improve your process.

### Peer Review

Peer review of preprint or publication using example template. The goal of this assignment is to focus on critiquing the logic put forward in the paper with suggestions for refining it.


## Semester project
Semester project outline of writing products.

### Project topic

Meeting with me or 2-6 sentence description of the research project you plan on doing this semester.

### Project outline

Short outline of full project highlighting the hypotheses to test, data to collect, and impact of the project. This includes no figures but could include a small table for organizing hypotheses.

### Project first draft

Full paper draft with figures. The important aspect of this submission is a first round of feedback for guiding conclusions and papers to read.

### Project peer review

Review of a peer's paper with the goal of improving the work by increasing clarity, readability and communication.

### Project final draft

Full, final project paper with figures, data, hypothesis, and interpretations. Focus is on clearly communicating the research work you have done.

### Project presentation

Presentation of your research project with a focus on interpretation of data collected or processed by you.

\newpage
<!-- The sections of code below that include "child=..." are drawing from seperate .md documents. These allow for updating text in a single spot for use in multiple syllabus documents. -->
```{r Academic disruption, child='general_components/academic_disruption.md'}

```

```{r Academic honesty, child='general_components/academic_honesty.md'}

```

## Course and Teacher evaluation
Student input for the purpose of course improvement is taken very seriously and will potentially be done periodically. Please take the time to evaluate this course and the instructor, especially at the end of the semester. Evaluations will in no way affect your grade.

\newpage
# Safety

## Field Safety

### Transportation
Wear your seatbelts at all times while in moving vehicles. Do not distract the driver.

### Location awareness
Be aware of potential dangers (animals, weather, etc) and avoid them. When in a group, please think about the safety of your peers.

## Building Safety

### Fire
Random fire safety drills may occur during the course of the term. You will be expected to evacuate the building and assemble at the designated location: outside by the northeast corner of the building (follow your instructor).

### Tornado
Move away from external windows to the hallway in the basement, first, or second floor. 

### Active Shooter Incident
"Get Out" of the building as long as they can do so safely. Once you have exited the building and are at a safe location, "Call Out" to law enforcement using 911, USA Police at 460-6312, or press the Emergency button on your LiveSafe Mobile Safety App. If you can't get out, "Shelter in Place" in a secure location that is away from windows and doors, preferably a location that can be locked and barricaded.

\newpage
# On writing

## Revising written work

Nothing is perfect on the first draft. We all revise and rework our written creations for a variety of reasons. We may not communicate the point we want to get across to all readers. We may have forgotten or not known important pieces of information initially. The tone of the writing may be inappropriate for the audience. Word choice might obscure meaning. The only way we can make our writing better is by revising it based on constructive feedback from others or ourselves.

## Genres of Science writing

### Paper reviews

Critical but constructive critique of unpublished research papers. Informs the journal editor of how the work contributes to existing literature and provides guide to authors and the editor for improvement to the manuscript. Can be signed but often anonymous.

### Descriptive Reports

Descriptive reports are often made for industry or geological surveys. These describe the composition and structure of rocks in an area. Sometimes they include estimations of resource potentials or environmental impacts of industry.

### Research papers

These pieces of science writing typically test one or more closely related hypotheses in the context of existing scientific knowledge. Most have structure that is pre-defined including: abstract, introduction, methods, results, discussion, and conclusion. Figures are an integral part of most papers in the Geosciences and range from images of outcrops to abstracted summaries of observations to plots of measurements.

### Conference abstracts

These are generally a step on the path to a full research publication where results, interpretations, and context are publicly presented for the first time.

### Grant Applications

Scientific work is generally funded with public or private funds that are obtained by writing grants. Unlike research papers, grants must persuade reviewers that the research planned is important, can be done by the applicant, and can be achieved in the time frame of the funding request.

### Science Communication

Scientists also write general summaries of their work for non-specialist audiences. These can be blog posts, or lesson plans that are shared for educators. Some folks also write books for general audiences describing the processes of science.

## Common processes

For writing in the discipline, we focus on several specific questions to guide our composition:
1) Who is our audience?
2) What is the one or two sentence take-home point?
3) How quickly are we communicating the information?

Science writing, like all writing, is socially situated and caries with it the beliefs, values, and ideologies of the particular community and culture.

## Tips and tricks for good writing
Use a citation management software to organize your papers and create bibiliographies. My preference is [Zotero](https://www.zotero.org/) because it is free, open source, and works across multiple operating systems.

Write and then revise. Ask for feedback from peers or wait until you have some distance from your writing and then revise.

Prompt yourself with questions that have short, concrete answers.

\newpage
# Campus resources

```{r COVID policy, child='general_components/COVID_policy.md'}

```

```{r Writing Center, child='general_components/writing_center.md'}

```

```{r Disability Services, child='general_components/disability_services.md'}

```

```{r Food Insecurity, child='general_components/food_insecurity.md'}

```

```{r Mental Health, child='general_components/mental_health.md'}

```

```{r Disaster Plan, child='general_components/disaster_plan.md'}

```

# Final thoughts

This document is a roadmap for our semester. We learn about the Earth together and our individual experiences shape how we interpret and value data. Like all your classes, you will get out what you put into this course. Asking for help from one another and your instructors is important, don't be afraid to ask a question about something you don't know or if you want to check your knowledge about something you think you know.

__If this document is updated, a copy will be supplied to you via Canvas and changes will be announced in class.__