---
title: 'GY111: Physical Geology'
date: "Spring 2022"
output:
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    latex_engine: pdflatex
    template: Syllabus_template.tex
  html_document:
    df_print: paged
geometry: margin=1in

#Fill this in with any number of instructors. It uses a two column layout with a column break after each instructor.

instructor:
- name: Dr. Benjamin J. Linzmeier
  email: blinzmeier@southalabama.edu
  prefered_contact: Canvas LMS messages
  officehours: 1:00 - 3:00 PM Mo
               3:00 - 5:00 PM Tu
               10:30 AM - 12:30 PM Fr
  office: Life Science Building room 344
  correspondence_policy: Within 24 hours on weekdays; 48 hours on weekends

# You can include multiple entries for a teaching assistant by filling in all these options below. Currently this is set to populate two columns on the syllabus but there is not an even column break inserted.

# teachingassistant:
#  - name: Filler
#    email: filler@fill.com
#    officehours: W T
#    office: OfficeNumber
#    sections: '301'
   
classroom: 'Mitchell Center Room 1101'
classhours: 9:05 AM - 9:55 AM	Mo We Fr
#labroom: 'LCSB 337'
#labhours: 1:25 pm - 3:55 pm	W
fontfamily: mathpazo
fontsize: 11pt
header-includes: \linespread{1.05}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(timeDate)
library(tibble)
library(dplyr)
library(knitr)

#This code chunk sets up a table of meeting days for the course.
#You can set days of the week for course meetings (e.g. Monday and Wednesday).
#Semester start and end dates are selected and then university holidays like spring break are specified.

#Days of the week that the course meets. Must be full day name in quotes.
Days <- c("Monday", "Wednesday", "Friday")

#Semester start date for classes as YYYY/M/D.
SemesterStart <- as.Date("2022/1/10")

#Semester end date for classes as YYYY/M/D.
SemesterEnd <- as.Date("2022/4/29")

#Fall break start and stop date.
FallBreak <-
  seq.Date(
    from = as.Date("2021/10/7"),
    to = as.Date("2021/10/10"),
    by = "day"
  )

#Spring break start and stop date.
SpringBreak <-
  seq.Date(
    from = as.Date("2022/02/28"),
    to = as.Date("2022/03/06"),
    by = "day"
  )

#Creates a list of days within the semester.
SemesterDays <-
  seq.Date(from = SemesterStart, to = SemesterEnd, by = "day")

#Add day before Thanksgiving.
DayBeforeThnks<-as.Date(holiday(
    year = getRmetricsOptions("currentYear"),
    Holiday = c("USThanksgivingDay")))-1

#Full list of holidays that there are no classes including breaks.
Holidays <-
  c(as.Date(holiday(
    year = getRmetricsOptions("currentYear"),
    Holiday = c("USLaborDay", "USThanksgivingDay", "USMLKingsBirthday")
  )), SpringBreak, FallBreak, DayBeforeThnks)

#Create a list of semester days excluding holidays.
SemesterDays <- SemesterDays[!SemesterDays %in% Holidays]

#Make dataframe with all semester days.
ClassDays <- data.frame(index = 1:length(SemesterDays))

#Insert dates.
ClassDays$Date <- SemesterDays

#Insert weekday designation.
ClassDays$Weekday <- weekdays(ClassDays$Date)

#Limit to the class meeting days.
MeetingDays <- ClassDays[ClassDays$Weekday %in% Days, ]

#Create numeric list of weeks.
Week <- cut.Date(MeetingDays$Date, breaks = "week")
levels(Week) <- 1:length(levels(Week))

MeetingDays$Week <- Week
MeetingDays$Class <- 1:nrow(MeetingDays)

#Replace this with c("Lecture topic 1",...,"Lecture topic N") to fill in table.

Lecture_Topic <- 1:nrow(MeetingDays)

Lecture_Topic <- c("Introduction and overview",
                   "Geological time",
                   "Discovery of geological time",
                   "Continental drift",
                   "Plate tectonics",
                   "Earthquakes",
                   "Minerals",
                   "Minerals",
                   "Mineral resources",
                   "The Rock cycle",
                   "Igneous Rocks",
                   "Igneous Rocks",
                   "Volcanoes",
                   "Volcanoes",
                   "Volcanic hazards",
                   "Metamorphism",
                   "Deformation and Landscapes",
                   "Mountains",
                   "Weathering",
                   "Erosion",
                   "Sedimentary Rocks",
                   "Planetary Geology",
                   "Geobiology",
                   "Geobiology",
                   "Wind",
                   "Deserts",
                   "Glaciers",
                   "Hydrologic Cycle",
                   "Groundwater",
                   "Streams",
                   "Flooding",
                   "Coastlines & Oceans",
                   "Climate System",
                   "Human impacts on climate",
                   "Review and synthesis",
                   "",
                   "",
                   "",
                   "",
                   "",
                   "",
                   "",
                   "",
                   "",
                   ""
                   )


MeetingDays <- bind_cols(MeetingDays, Lecture = Lecture_Topic)

#### Grade table ####
Grades <- tribble( ~ Grade,
                   ~ Range,
                   "A",
                   "90-100",
                   "B",
                   "80-89",
                   "C",
                   "70-79",
                   "D",
                   "60-69",
                   "F",
                   "< 60")

####Assignment weight table####
Weights <- tribble(
                    ~ Item,
                    ~ Weight,
                    "Discussion posts",
                    "~12.5 %",
                    "In-class work",
                    "~12.5 %",
                    "Readings",
                    "~42 %",
                    "Quizzes",
                    "~16 %",
                    "Final Exam",
                    "~16 %"
                  )


```

# Course description

How does the Earth work as a system? In this class we will learn the basics of the science of Geology with a focus on physical processes and materials. We will cover major topics including: plate tectonics, mineral chemistry, the rock cycle, geological time, sedimentary processes, climate change, and geological map reading.

# Important Dates
January 17th - Last day to drop without WD\
April 1st - Last day to drop or withdraw\
See the [Academic Calendar](https://www.southalabama.edu/academiccalendar/) for more information.

# Course goals

1.    Explain and apply the scientific method as it is practiced in the Geosciences
2.    Link observations of the Earth on multiple scales to the development of theories
3.    Explain how the hydrosphere, biosphere, and geosphere interact on multiple timescales and predict how changes in one can affect others
4.    Read and interpret topographic and geological maps
5.    Identify common minerals and rocks
6.    Link geological materials to Earth processes
7.    Relate facts about the Earth system to human problems


# Class textbook

We will be using the e-book of Exploring Geology by Stephen Reynolds and Julia Johnson. I chose this book because it has guided practice with readings and includes interactive web-based activities based on societally relevant problems related to Geology.


### Lectures and papers
Both lecture slides and supplemental readings will be available via Canvas. Slides will generally be posted after lecture.


\newpage

```{r Schedule, echo=FALSE}
kable(MeetingDays[,c("Date",
                     "Class",
                     "Week",
                     "Lecture")
                     ],
  align = c("l",
            "c",
            "c",
            "l")
)
```

\newpage

# Student conduct

## Attendance policy

-   Students who complete less than 50% of assignments/exams will be assigned a failing (F) grade.

-   Review the 'Attendance and Absences Policy', in the Undergraduate and Graduate Bulletin for attendance and absences policy

## Grading

### Missing work
Everyone will have their lowest two discussion posts, readings, and quizzes dropped. This means that you have two weeks of built-in flexibility for life events that come up (e.g. illness, unexpected travel).

### Late work policy
If you need additional flexibility based on difficulties arising from life, please let me know. That said, given dropping the lowest two weeks worth of work means that I will not be allowing for late quizzes or discussion posts.

### Extra credit
Extra credit may be made available throughout the semester, but readings, quizzes, and discussion posts .

Assignments will be returned within a week of submission.

```{r Grades_Table, echo=FALSE}
kable(Grades,
      align = c("l", "r"))
```

Assignment weighting will follow:
```{r Weights_Table, echo=FALSE}
kable(Weights,
      align = c("l", "r"))
```

<!-- The sections of code below that include "child=..." are drawing from seperate .md documents. These allow for updating text in a single spot for use in multiple syllabus documents. -->
```{r Academic disruption, child='general_components/academic_disruption.md'}

```

```{r Academic honesty, child='general_components/academic_honesty.md'}

```

## Course and Teacher evaluation
Student input for the purpose of course improvement is taken very seriously and will potentially be done periodically. Please take the time to evaluate this course and the instructor, especially at the end of the semester. Evaluations will in no way affect your grade.

\newpage
# Safety

## Building Safety

### Fire
Random fire safety drills may occur during the course of the term. You will be expected to evacuate the building and assemble at the designated location: outside by the northeast corner of the building (follow your instructor).

### Tornado
Move away from external windows to the hallway in the basement, first, or second floor. 

### Active Shooter Incident
"Get Out" of the building as long as they can do so safely. Once you have exited the building and are at a safe location, "Call Out" to law enforcement using 911, USA Police at 460-6312, or press the Emergency button on your LiveSafe Mobile Safety App. If you can't get out, "Shelter in Place" in a secure location that is away from windows and doors, preferably a location that can be locked and barricaded.

\newpage
# Campus resources

```{r COVID policy, child='general_components/COVID_policy.md'}

```

```{r Writing Center, child='general_components/writing_center.md'}

```

```{r Disability Services, child='general_components/disability_services.md'}

```

```{r Food Insecurity, child='general_components/food_insecurity.md'}

```

```{r Mental Health, child='general_components/mental_health.md'}

```

```{r Disaster Plan, child='general_components/disaster_plan.md'}

```

# Final thoughts

This document is a roadmap for our semester. We learn about the Earth together and our individual experiences shape how we interpret and value data. Like all your classes, you will get out what you put into this course. Asking for help from one another and your instructors is important, don't be afraid to ask a question about something you don't know or if you want to check your knowledge about something you think you know.

__If this document is updated, a copy will be supplied to you via Canvas and changes will be announced in class.__